---
title: "Overall Responses"
author: "Amanda Skarlupka"
date: "10/27/2019"
output: html_document
---
This document contains Figures 3 and 5 and supplemental figures 1 through 4. 
It investigates the overall inhibition of P1, Wildtype, and CA/09 monoclonal antibodies. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


Load the needed packages.
```{r}
library(ggplot2)
library(forcats)
library(broom)
library(ggbiplot)
library(dplyr)
library(here)
```


Load the processed data
```{r}
ca09_mab <- readRDS(here::here("data", "processed_data", "ca09mab_processed_data.rds"))
ca09_sera <- readRDS(here::here("data", "processed_data", "ca09sera_processed_data.rds"))
p1_mab <- readRDS(here::here("data", "processed_data", "p1mab_processed_data.rds"))
p1_sera <- readRDS(here::here("data", "processed_data", "p1sera_processed_data.rds"))
```


take a look at the data
```{r}
glimpse(ca09_mab)
glimpse(ca09_sera)
glimpse(p1_mab)
glimpse(p1_sera)
```


Example graph to try and get the range of reaction of the monoclonal antibodies. Use the 1F8 as an example because it's supposed to be the most interesting one due to broad reactivity. 
```{r}
p1_mab$lineage[is.na(p1_mab$lineage)] <- "unknown"
  
p1_mab$lineage <- factor(p1_mab$lineage, levels = c("human_like", "unknown", "classical", "eurasian"))
p1_mab$sublineage <- factor(p1_mab$sublineage, levels = c("alpha", "beta", "gamma", "gamma_2", "npdm", "delta_1", "delta_2"))
p1_mab$antigen <- factor(p1_mab$antigen, levels = c("Swine/31", "Iowa/73", "WI/97", "Indiana/00", "NC/01", "Spain/2003", "Zhejiang/07", "Colorado/09", "Illinois/09", "Minnesota/09", "NC/09", "NC/34543/09", "NC/5043-1/09", "Utah/09", "Missouri/13", "Nebraska/13", "MN/15", "NC/15"), ordered = TRUE)
f <- p1_mab %>%
  filter(clone_name == "1F8") %>%
  select(c(dilution, antigen, sublineage, lineage, clone_name)) %>%
  arrange(lineage, sublineage, antigen) %>%
  pull(antigen)

p1_mab$antigen <- factor(p1_mab$antigen, levels = (f), ordered=TRUE)

#testing the code
p1_mab %>%
  filter(clone_name == "1F8") %>%
  ggplot(aes(y = dilution, x = antigen)) +
  geom_point(aes(shape = lineage, color = sublineage, size = 3)) +
  theme(axis.text.x = element_text(angle = 60, hjust = 1)) +
  scale_shape_manual(values = c(15:18)) +
  scale_y_continuous(trans = "log2") +
  labs(y = "Dilution from 20ug/ml") 

```



Graph of all the monoclonal responses

```{r}
p1mab_by_mab <- p1_mab %>%
  ggplot(aes(y = dilution, x = antigen, shape = lineage, color = sublineage), size = 0.5) +
  geom_point() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  labs(y = "Dilution from 20ug/ml",
       x = "Antigen") +
  scale_shape_manual(values = c(15:18)) +
  scale_y_continuous(trans = "log2") +
  facet_wrap(vars(clone_name))

pu <- p1_mab %>%
  arrange(elisa_specificity) %>%
  pull(clone_name)
pu <- unique(pu)
p1_mab$clone_name <- factor(p1_mab$clone_name, levels = pu, ordered = TRUE)

p1mab_by_antigen <- p1_mab %>%
  ggplot(aes(y = dilution, x = clone_name, color = elisa_specificity), size = 0.5) +
  geom_point() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  labs(y = "Dilution from 20ug/ml",
       x = "Antigen") +
  scale_shape_manual(values = c(15:18)) +
  scale_y_continuous(trans = "log2") +
  facet_wrap(vars(antigen))

ggsave(filename= here("results", "figures", "p1mab_by_mab.png"), plot = p1mab_by_mab)
p1mab_by_mab
ggsave(filename = here("results", "figures", "supp_p1mab_by_antigen.png"), plot = p1mab_by_antigen)
p1mab_by_antigen
```
Repeat it for the CA09 specific monoclonals
```{r}
ca09_mab$lineage[is.na(ca09_mab$lineage)] <- "unknown"
  
ca09_mab$lineage <- factor(ca09_mab$lineage, levels = c("human_like", "unknown", "classical", "eurasian"))
ca09_mab$sublineage <- factor(ca09_mab$sublineage, levels = c("alpha", "beta", "gamma", "gamma_2", "npdm", "delta_1", "delta_2"))
ca09_mab$antigen <- factor(ca09_mab$antigen, levels = c("Swine/31", "Iowa/73", "WI/97", "Indiana/00", "NC/01", "Spain/2003", "Zhejiang/07", "Colorado/09", "Illinois/09", "Minnesota/09", "NC/09", "NC/34543/09", "NC/5043-1/09", "Utah/09", "Missouri/13", "Nebraska/13", "MN/15", "NC/15"), ordered = TRUE)

ca09_mab$antigen <- factor(ca09_mab$antigen, levels = (f), ordered=TRUE)

#testing the code
ca09_mab %>%
  filter(clone_name == "1E6") %>%
  ggplot(aes(y = dilution, x = antigen)) +
  geom_point(aes(shape = lineage, color = sublineage, size = 3)) +
  theme(axis.text.x = element_text(angle = 60, hjust = 1)) +
  scale_shape_manual(values = c(15:18)) +
  labs(y = "Dilution from 20ug/ml") 
```
We'll now look at the wild-type raised monoclonals. 
```{r}

cn <- ca09_mab %>%
  arrange(raised_against) %>%
  pull(clone_name)
cn <- unique(cn)
ca09_mab$clone_name <- factor(ca09_mab$clone_name, levels = (cn), ordered=TRUE)

WTmab_by_mab <- ca09_mab %>%
  ggplot(aes(y = dilution, x = antigen, shape = lineage, color = sublineage), size = 0.5) +
  geom_point() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  labs(y = "Dilution from 20ug/ml") +
  scale_shape_manual(values = c(15:18)) +
  scale_y_continuous(trans = "log2") +
  facet_wrap(vars(clone_name))

WTmab_by_antigen <- ca09_mab %>%
  ggplot(aes(y = dilution, x = clone_name, shape = elisa_specificity, color = raised_against), size = 0.5) +
  geom_point() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  labs(y = "Dilution from 20ug/ml",
       x = "Antigen") +
  scale_shape_manual(values = c(15:18)) +
  scale_y_continuous(trans = "log2") +
  facet_wrap(vars(antigen))

WTmab_by_antigen
WTmab_by_mab

ggsave(filename = here("results", "figures", "supp_WTmab_by_mab.png"), plot = WTmab_by_mab, width = 9)
ggsave(filename = here("results", "figures", "supp_WTmab_by_antigen.png"), plot = WTmab_by_antigen, width = 9)
```


The previous graphs included the all of the wild-type antibodies (including SC/18 and Bris/07). We're specifically interested in CA/09. So I'll make a graph of just CA/09 and save the other as a supplemental image. 

```{r}
ca09_mab$key_name <- as.factor(ca09_mab$key_name)

ca09_mab_individual <- ca09_mab %>%
  filter(raised_against == "ca09") %>%
  ggplot(aes(y = dilution, x = antigen, color = lineage), size = 0.5) +
  geom_point() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  labs(y = "Dilution from 20ug/ml") +
  scale_shape_manual(values = c(15:18)) +
  scale_y_continuous(trans = "log2") +
  facet_wrap(vars(clone_name))

ca09_mab_by_antigen <- ca09_mab %>%
  filter(raised_against == "ca09") %>%
  ggplot(aes(y = dilution, x = clone_name), size = 0.5) +
  geom_point() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  labs(y = "Dilution from 20ug/ml",
       x = "Antigen") +
  scale_shape_manual(values = c(15:18)) +
  scale_y_continuous(trans = "log2") +
  facet_wrap(vars(antigen))

ggsave(here("results", "figures", "ca09_mab_individual.png"), plot = ca09_mab_individual, width = 9)
ca09_mab_individual

ggsave(here("results", "figures", "supp_ca09_mab_by_antigen.png"), plot = ca09_mab_by_antigen, width =9)
```



  