---
title: "Overall Responses"
author: "Amanda Skarlupka"
date: "10/27/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


Load the needed packages.
```{r}
library(ggplot2)
library(forcats)
library(broom)
library(ggbiplot)
library(dplyr)
library(here)
```


Load the processed data
```{r}
ca09_mab <- readRDS(here("data", "processed_data", "ca09mab_processed_data.rds"))
ca09_sera <- readRDS(here("data", "processed_data", "ca09sera_processed_data.rds"))
p1_mab <- readRDS(here("data", "processed_data", "p1mab_processed_data.rds"))
p1_sera <- readRDS(here("data", "processed_data", "p1sera_processed_data.rds"))
```



#I need these because I wrote the code before I realized that my data wasn't actually linked. So I'll need to go back and clean this up.
CA09_mabs_full <- readRDS("./data/processed_data/CA09mabs_full_processed_data.rds")
CA09_sera_full <- readRDS("./data/processed_data/CA09sera_full_processed_data.rds")
p1_mabs_full <- readRDS("./data/processed_data/p1mabs_full_processed_data.rds")
p1_sera_full <- readRDS("./data/processed_data/p1sera_full_processed_data.rds")

take a look at the data
```{r}
glimpse(ca09_mab)
glimpse(ca09_sera)
glimpse(p1_mab)
glimpse(p1_sera)
```


Example graph to try and get the range of reaction of the monoclonal antibodies. Use the 1F8 as an example because it's supposed to be the most interesting one due to broad reactivity. 
```{r}
p1_mab$lineage[is.na(p1_mab$lineage)] <- "unknown"
  
p1_mab$lineage <- factor(p1_mab$lineage, levels = c("human_like", "unknown", "classical", "eurasian"))
p1_mab$sublineage <- factor(p1_mab$sublineage, levels = c("alpha", "beta", "gamma", "gamma_2", "npdm", "delta_1", "delta_2"))
p1_mab$antigen <- factor(p1_mab$antigen, levels = c("Swine/31", "Iowa/73", "WI/97", "Indiana/00", "NC/01", "Spain/2003", "Zhejiang/07", "Colorado/09", "Illinois/09", "Minnesota/09", "NC/09", "NC/34543/09", "NC/5043-1/09", "Utah/09", "Missouri/13", "Nebraska/13", "MN/15", "NC/15"), ordered = TRUE)
f <- p1_mab %>%
  filter(clone_name == "1F8") %>%
  select(c(dilution, antigen, sublineage, lineage, clone_name)) %>%
  arrange(lineage, sublineage, antigen) %>%
  pull(antigen)

p1_mab$antigen <- factor(p1_mab$antigen, levels = (f), ordered=TRUE)

p1_mab %>%
  filter(clone_name == "1F8") %>%
  ggplot(aes(y = dilution, x = antigen)) +
  geom_point(aes(shape = lineage, color = sublineage, size = 3)) +
  theme(axis.text.x = element_text(angle = 60, hjust = 1)) +
  scale_shape_manual(values = c(15:18)) +
  scale_y_continuous(trans = "log2") +
  labs(y = "Dilution from 20ug/ml") 

```



Graph of all the monoclonal responses

```{r}
p1mab_by_mab <- p1_mab %>%
  ggplot(aes(y = dilution, x = antigen, shape = lineage, color = sublineage), size = 0.5) +
  geom_point() +
  theme(axis.text.x = element_text(angle = 60, hjust = 1)) +
  labs(y = "Dilution from 20ug/ml") +
  scale_shape_manual(values = c(15:18)) +
  scale_y_continuous(trans = "log2") +
  facet_wrap(vars(clone_name))

pu <- p1_mab %>%
  arrange(elisa_specificity) %>%
  pull(clone_name)
pu <- unique(pu)
p1_mab$clone_name <- factor(p1_mab$clone_name, levels = pu, ordered = TRUE)

p1mab_by_antigen <- p1_mab %>%
  ggplot(aes(y = dilution, x = clone_name, color = elisa_specificity), size = 0.5) +
  geom_point() +
  theme(axis.text.x = element_text(angle = 60, hjust = 1)) +
  labs(y = "Dilution from 20ug/ml") +
  scale_shape_manual(values = c(15:18)) +
  scale_y_continuous(trans = "log2") +
  facet_wrap(vars(antigen))

ggsave(filename= here("results", "p1mab_by_mab.png"), plot = p1mab_by_mab)
ggsave(filename = here("results", "p1mab_by_antigen.png"), plot = p1mab_by_antigen)
```
Repeat it for the CA09 specific monoclonals
```{r}
ca09_mab$lineage[is.na(ca09_mab$lineage)] <- "unknown"
  
ca09_mab$lineage <- factor(ca09_mab$lineage, levels = c("human_like", "unknown", "classical", "eurasian"))
ca09_mab$sublineage <- factor(ca09_mab$sublineage, levels = c("alpha", "beta", "gamma", "gamma_2", "npdm", "delta_1", "delta_2"))
ca09_mab$antigen <- factor(ca09_mab$antigen, levels = c("Swine/31", "Iowa/73", "WI/97", "Indiana/00", "NC/01", "Spain/2003", "Zhejiang/07", "Colorado/09", "Illinois/09", "Minnesota/09", "NC/09", "NC/34543/09", "NC/5043-1/09", "Utah/09", "Missouri/13", "Nebraska/13", "MN/15", "NC/15"), ordered = TRUE)

ca09_mab$antigen <- factor(ca09_mab$antigen, levels = (f), ordered=TRUE)

ca09_mab %>%
  filter(clone_name == "1E6") %>%
  ggplot(aes(y = dilution, x = antigen)) +
  geom_point(aes(shape = lineage, color = sublineage, size = 3)) +
  theme(axis.text.x = element_text(angle = 60, hjust = 1)) +
  scale_shape_manual(values = c(15:18)) +
  labs(y = "Dilution from 20ug/ml") 
```

```{r}

cn <- ca09_mab %>%
  arrange(raised_against) %>%
  pull(clone_name)
cn <- unique(cn)
ca09_mab$clone_name <- factor(ca09_mab$clone_name, levels = (cn), ordered=TRUE)

ca09mab_by_mab <- ca09_mab %>%
  ggplot(aes(y = dilution, x = antigen, shape = lineage, color = sublineage), size = 0.5) +
  geom_point() +
  theme(axis.text.x = element_text(angle = 60, hjust = 1)) +
  labs(y = "Dilution from 20ug/ml") +
  scale_shape_manual(values = c(15:18)) +
  scale_y_continuous(trans = "log2") +
  facet_wrap(vars(clone_name))

ca09mab_by_antigen <- ca09_mab %>%
  ggplot(aes(y = dilution, x = clone_name, shape = elisa_specificity, color = raised_against), size = 0.5) +
  geom_point() +
  theme(axis.text.x = element_text(angle = 60, hjust = 1)) +
  labs(y = "Dilution from 20ug/ml") +
  scale_shape_manual(values = c(15:18)) +
  scale_y_continuous(trans = "log2") +
  facet_wrap(vars(antigen))

ca09mab_by_antigen

ggsave(filename = here("results", "ca09mab_by_mab.png"), plot = ca09mab_by_mab)
ggsave(filename = here("results", "ca09mab_by_antigen.png"), plot = ca09mab_by_antigen)
```
potentially include the SW1-4 COBRA VLPs. 

Graph of just wild-type H1 raised monoclonal responses.

```{r}
ca09_mab$key_name <- as.factor(ca09_mab$key_name)
#ca09_mab$key_name <- mutate(ca09_mab$key_name, funs=toupper)



ca09_mab %>%
  filter(raised_against == "ca09") %>%
  ggplot(aes(y = dilution, x = antigen, color = lineage), size = 0.5) +
  geom_point() +
  theme(axis.text.x = element_text(angle = 60, hjust = 1)) +
  labs(y = "Dilution from 20ug/ml") +
  scale_shape_manual(values = c(15:18)) +
  scale_y_continuous(trans = "log2") +
  facet_wrap(vars(clone_name))


```



  