---
title: "CA/09 Mab Reactivity"
author: "Amanda Skarlupka"
date: "11/8/2019"
output: html_document
---

Figures 2 and 6 are contained in this document. 

This CA/09 specific document contains the radar/spider graphs of the HAI response. It also contains the HAI responses based on elisa specificity of the monoclonals. 
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

I need to load the packages. This'll follow a lot of the P1 mabs reactivity analysis. 

```{r}
library(ggplot2)
library(forcats)
library(broom)
library(ggbiplot)
library(dplyr)

library(ggradar)
library(tibble)
library(ggpubr)
library(scales)
library(here)
```

I need to load the data that has the wildtype reactivity

```{r}
ca09_mab <- readRDS(here::here("data", "processed_data", "ca09mab_processed_data.rds"))
ca09_sera <- readRDS(here::here("data", "processed_data", "ca09sera_processed_data.rds"))
antigen_key <- readRDS(here::here("data", "processed_data", "antigen_key.rds"))
p1_sera <- readRDS(here::here("data", "processed_data", "p1sera_processed_data.rds"))
p1_mab <- readRDS(here::here("data", "processed_data", "p1mab_processed_data.rds"))
```
Pan-H1 seasonal and pdm mAbs

* AT163.272.54
* 3B6

HA Stem control mAbs (more than likely not included due to no HAI reactivity)

* C179
* CR6261
* FI6
* F10

Narrowly-reactive seasonal mAbs

* AT170.558.146
* AT163.210.182
* AT163.104.93
* AT170.119.5
* AT163.333.93
* AT163.329.189

Narrowly-reactive 1918 pdm and 1934 mAbs

* 5D3
* 6B9
* 39E4
* IC5-4F8
* RA5-22

Narrowly-reactive CA/09 pdm mAbs

* 1C5
* 5C12
* CA09-02
* CA09-09
* CA09-11
* CA09-15
* 1E6
* 2A12
* 2B11
* 1G12
* 3G6
* 5B-2A12
* 2G7
* 2F3
* 3B8
* M2-2B9


```{r}
ca09_mab$lineage[is.na(ca09_mab$lineage)] <- "unknown"
  
ca09_mab$lineage <- factor(ca09_mab$lineage, levels = c("human_like", "unknown", "classical", "eurasian"))
ca09_mab$sublineage <- factor(ca09_mab$sublineage, levels = c("alpha", "beta", "gamma", "gamma_2", "npdm", "delta_1", "delta_2"))
ca09_mab$antigen <- factor(ca09_mab$antigen, levels = c("Swine/31", "Iowa/73", "WI/97", "Indiana/00", "NC/01", "Spain/2003", "Zhejiang/07", "Colorado/09", "Illinois/09", "Minnesota/09", "NC/09", "NC/34543/09", "NC/5043-1/09", "Utah/09", "Missouri/13", "Nebraska/13", "MN/15", "NC/15"), ordered = TRUE)
f <- ca09_mab %>%
  filter(clone_name == "1E6") %>%
  select(c(dilution, antigen, sublineage, lineage, clone_name)) %>%
  arrange(lineage, sublineage, antigen) %>%
  pull(antigen)

ca09_mab$antigen <- factor(ca09_mab$antigen, levels = (f), ordered=TRUE)

```

```{r}
ca09_mab_fig4 <- ca09_mab %>%
  ggplot(aes(y = dilution, x = antigen)) +
  geom_col(aes(shape = lineage, fill = clone_name, color = clone_name), position = "dodge") +
  theme(axis.text.x = element_text(angle = 60, hjust = 1)) +
  scale_shape_manual(values = c(15:18)) +
  scale_y_continuous(trans = "log2") +
  theme(axis.text.x = element_text(colour = c("blue", "blue", "blue", "blue", "grey", "red", "red", "red", "red", "red", "red", "red", "red", "red", "red", "red", "black", "black", "blue", "blue"))) +
  labs(y = "Dilution from 20ug/ml",
       title = "CA/09 Monoclonals against Swine Virus HAs") +
  facet_grid(rows= vars(elisa_specificity))
ggsave(filename = here::here("results", "figures", "ca09_mab_fig4.png"), plot = ca09_mab_fig4)

ca09_mab_fig4
```


Create a radar plot to try and visualize the differences of the broad, narrow, and cobra specific monoclonals in one plot. 

Change into log and reorder the columns so that they are sorted by the lineages of the swine HAs.
```{r}

ca09_mab_full <- readRDS(here::here("data", "processed_data", "ca09mab_full.rds"))
ca09_mab_full[, c(2:19)] <- log2(20/ca09_mab_full[,c(2:19)])

swine_lineage <- c("clone_name", "Swine/31", "Iowa/73", "WI/97", "Indiana/00", "NC/01", "Spain/2003", "Zhejiang/07", "Colorado/09", "Illinois/09", "Minnesota/09", "NC/09", "NC/34543/09", "NC/5043-1/09", "Utah/09", "Missouri/13", "Nebraska/13", "MN/15", "NC/15")

ca09_mab_full <- ca09_mab_full[, swine_lineage]

ca09_mab_full <- ca09_mab_full %>%
  select(-c("Missouri/13", "NC/15", "Illinois/09"))
  
```


Narrow reactivity to CA/09 mAbs
```{r}
narrow_ca09 <- ca09_mab %>%
  filter(elisa_specificity == "narrow_ca09") %>%
  pull(clone_name)
narrow_ca09 <- unique(narrow_ca09)

ca09_mab_narrow_ca09 <- ca09_mab_full %>%
  filter(clone_name %in% narrow_ca09)

radar_narrow_ca09 <- ggradar(ca09_mab_narrow_ca09, values.radar = c(0, 1, 2, 3, 4, 5, 6, 7, 8, 9, 10), grid.max = 10, group.point.size = 1, group.line.width = 0.5, legend.title = "Narrow reactivity to CA/09")
```

Narrow reactivity to seasonal mAbs
```{r}
narrow_seasonal <- ca09_mab %>%
  filter(elisa_specificity == "narrow_seasonal") %>%
  pull(clone_name)
narrow_seasonal <- unique(narrow_seasonal)

ca09_mab_narrow_seasonal <- ca09_mab_full %>%
  filter(clone_name %in% narrow_seasonal)

radar_narrow_seasonal <- ggradar(ca09_mab_narrow_seasonal, values.radar = c(0, 1, 2, 3, 4, 5, 6, 7, 8, 9, 10), grid.max = 10, group.point.size = 1, group.line.width = 0.5, legend.title = "Narrow reactivity to seasonal H1s")

```

Narrow reactivity to 1918 pandemic strains
```{r}
narrow_1918pdm <- ca09_mab %>%
  filter(elisa_specificity == "narrow_1918pdm") %>%
  pull(clone_name)
narrow_1918pdm <- unique(narrow_1918pdm)

ca09_mab_narrow_1918pdm <- ca09_mab_full %>%
  filter(clone_name %in% narrow_1918pdm)

radar_narrow_1918pdm <- ggradar(ca09_mab_narrow_1918pdm, values.radar = c(0, 1, 2, 3, 4, 5, 6, 7, 8, 9, 10), grid.max = 10, group.point.size = 1, group.line.width = 0.5, legend.title = c("Narrow reactivity to seasonal H1s"))

```

Plot all of them on the same figure
```{r}

radar_fig5 <- ggarrange(radar_narrow_ca09, radar_narrow_seasonal, radar_narrow_1918pdm, ncol=3, nrow = 1)

ggsave(filename = here::here("results", "figures", "radar_fig5.png"), plot = radar_fig5, width = 8, height = 4, dpi = 600, scale = 4)

```



