---
title: "PCA Analysis"
author: "Amanda Skarlupka"
date: "10/30/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(ggplot2)
library(forcats)
library(broom)
library(ggbiplot)
library(dplyr)
library(here)
```


Load the processed data
```{r}

ca09_mab_full <- readRDS(here("data", "processed_data", "ca09mab_full.rds"))
ca09_sera_full <- readRDS(here("data", "processed_data", "ca09sera_full.rds"))
p1_mab_full <- readRDS(here("data", "processed_data", "p1mab_full.rds"))
p1_sera_full <- readRDS(here("data", "processed_data", "p1sera_full.rds"))
antibody_key <- readRDS(here("data", "processed_data", "antibody_key.rds"))

```

Principal component analysis on antibody responses to different viruses
Need to remove all the mice that do not have full matrix

```{r}
p1_sera <- p1_sera_full %>%
  select(-c(4:7, 13:14, 16)) %>%
  filter(mouse != 2 & mouse != 10)
```

Need to remove mice that do not have variability for a virus. 
```{r}
p1_sera <- p1_sera[,apply(p1_sera, 2, var, na.rm=TRUE) != 0]
```

Do pca analysis
```{r}
p1_sera_pca <- prcomp(p1_sera[,2:8], center = TRUE, scale = TRUE)
summary(p1_sera_pca)
str(p1_sera_pca)
p1_sera_var <- plot(p1_sera_pca, type = "l")

#ggsave(filename = here("results", "p1sera_pca_var.png"), plot = p1_sera_var)

g <- ggbiplot(p1_sera_pca, obs.scale = 1, var.scale = 1, groups = p1_sera$mouse)
g
```

The sera dilution is in 2-fold dilutions. Therefore the data should be transformed to account for this, and the PCA reran. These analyses are saying that there are different clusters of antigens when mice are vaccinated with p1 vaccine. 
```{r}
p1_sera[1:8, 2:8] <- log2(p1_sera[1:8, 2:8])
p1_sera_pca <- prcomp(p1_sera[,2:8], center = TRUE, scale = TRUE)
ggbiplot(p1_sera_pca, obs.scale = 1, var.scale = 1, groups = p1_sera$mouse)
```



Repeat with CA09_sera
```{r}
ca09_sera <- ca09_sera_full %>%
  select(-c(4:7, 13:14, 16))
```

Remove all the mice that do not have full matrix and all antigens that do not have variability to the vaccine used. 
```{r}
ca09_sera <- ca09_sera[,apply(ca09_sera, 2, var, na.rm=TRUE) != 0]
ca09_sera_pca <- prcomp(ca09_sera[,2:6], center = TRUE, scale = TRUE)
summary(ca09_sera_pca)
str(ca09_sera_pca)
ca09_sera_var <- plot(ca09_sera_pca, type = "l")
ggbiplot(ca09_sera_pca, obs.scale = 1, var.scale = 1, groups = ca09_sera$mouse)
```

Do pca analysis

```{r}
ca09_sera[1:6, 2:6] <- log2(ca09_sera[1:6, 2:6])
ca09_sera_pca <- prcomp(ca09_sera[,2:6], center = TRUE, scale = TRUE)
ca09_sera_var <- plot(ca09_sera_pca, type = "l")
ggbiplot(ca09_sera_pca, obs.scale = 1, var.scale = 1, groups = ca09_sera$mouse)

```

There's not much variation, but if we add together the CA09_sera maybe we can see a little more variation
Renumber p1 mice so they don't overlap with the CA09 mice ids
```{r}
ca09_sera <- ca09_sera_full %>%
  select(-c(4:7, 13:14, 16))
p1_sera <- p1_sera_full %>%
  select(-c(4:7, 13:14, 16)) %>%
  filter(mouse != 2 & mouse != 10)
p1_sera$mouse <- c(7:14)

sera <- bind_rows("ca09" = ca09_sera, "p1" = p1_sera, .id = "vaccine")

sera <- sera %>%
  select(-c(9:12))

sera[1:14, 3:9] <- log2(sera[1:14, 3:9])

sera_pca<- prcomp(sera[,3:9], center = TRUE, scale = TRUE)
plot(sera_pca, type = "l")
ggbiplot(sera_pca, obs.scale = 1, var.scale = 1, groups = sera$vaccine, labels = sera$mouse) +
  labs(title = "PCA of CA/09 and P1 Sera to swH1N1/N2 viruses") +
  xlim(c(3.5, -3.5)) + 
  ylim(c(-3.5, 3.5))
```

The plot above is also known as variable correlation plots. It shows the relationships between all variables. It can be interpreted as follow:

Positively correlated variables are grouped together.
Negatively correlated variables are positioned on opposite sides of the plot origin (opposed quadrants).
The distance between variables and the origin measures the quality of the variables on the factor map. Variables that are away from the origin are well represented on the factor map.


Repeat for the CA09-raised monoclonals
```{r}
ca09_mab <- ca09_mab_full
ca09_mab <- ca09_mab[, apply(ca09_mab, 2, var, na.rm = TRUE) != 0]
ca09_mab[1:18, 2:14] <- log2(ca09_mab[1:18, 2:14])

ca09_mab_pca <- prcomp(ca09_mab[, 2:14], center = TRUE, scale = TRUE)
plot(ca09_mab_pca, type = "l")
ggbiplot(ca09_mab_pca, obs.scale = 1, var.scale = 1, groups = ca09_mab$clone_name)
```


Repeat for the P1-raised monoclonals
```{r}
p1_mab <- p1_mab_full %>%
  select(-c(5, 13, 15:16, 18))
p1_mab[1:12, 2:14] <- log2(p1_mab[1:12, 2:14])

p1_mab_pca <- prcomp(p1_mab[, 2:14], center = TRUE, scale = TRUE)
plot(p1_mab_pca, type = "l")
ggbiplot(p1_mab_pca, obs.scale = 1, var.scale = 1, groups = p1_mab$clone_name)
```


Put all monoclonals together
```{r}

mab <- bind_rows("ca09" = ca09_mab_full, "p1" = p1_mab_full, .id = "group")
mab <- mab %>%
  select(-c(6, 16:17, 19))
mab[1:30, 3:16] <- log2(mab[1:30, 3:16])

mab_pca <- prcomp(mab[, 3:16], center = TRUE, scale = TRUE)

mab <- antibody_key %>%
  select(c(clone_name, raised_against, elisa_specificity)) %>%
  right_join(mab, by = "clone_name")
plot(mab_pca, type = "l")
ggbiplot(mab_pca, obs.scale = 1, var.scale = 1, groups = mab$raised_against, var.axes = FALSE, alpha = 0.5)
```


```{r}
ca09_sera_mean <- data.frame(lapply(sera[1:6, 3:9], mean))

p1_sera_mean <- data.frame(lapply(sera[7:14, 3:9], mean))

t <- union(ca09_sera_mean, p1_sera_mean)

t %>%
  as_tibble() %>%
  mutate(group = c("ca09_sera", "p1_sera"))

```


