---
title: "Reactivity Analysis"
author: "Amanda Skarlupka"
date: "11/21/2019"
output: html_document
---
This document contains the code for Figure 8, supplementary figure 5 
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load the needed packages.
```{r}
library(ggplot2)
library(forcats)
library(broom)
library(ggbiplot)
library(dplyr)
library(here)
```


Load the processed data
```{r}
ca09_mab <- readRDS(here::here("data", "processed_data", "ca09mab_processed_data.rds"))
ca09_sera <- readRDS(here::here("data", "processed_data", "ca09sera_processed_data.rds"))
p1_mab <- readRDS(here::here("data", "processed_data", "p1mab_processed_data.rds"))
p1_sera <- readRDS(here::here("data", "processed_data", "p1sera_processed_data.rds"))
```

The questions that are being pursued here are:
1. What are the total number of viruses that each monoclonal inhibits the HA of?

A concentration of greater than 20, so just being recognized by the monoclonal antibody is what I'm considering to be inhibitory. The physiological relevant concentration is not known by me however.

```{r}
mabs <- bind_rows(ca09_mab, p1_mab)

univariate_clone <- mabs %>%
  filter(concentration < 20) %>%
  ggplot(aes(x = forcats::fct_infreq(clone_name))) +
  geom_bar(aes(fill = raised_against))+
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  labs(title = "Total number of viruses recognized above the limit of detection",
       x = "Monoclonal Antibody",
       y = "Number of Viruses")

ggsave(filename = here::here("results", "figures", "univariate_clone.png"), plot = univariate_clone)

univariate_clone

```
The CA09-09 monoclonal antibody recognized the most antigens and also recognized the most narrow selection of antigens as well. (2A12, 2B11, and CA09-11). 


2. How many monoclonals inhibit each virus?

```{r}
univariate_antigen <- mabs %>%
  filter(concentration < 20) %>%
  ggplot(aes(x = forcats::fct_infreq(antigen))) +
  geom_bar(aes(fill = raised_against)) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, colour = c("red", "red", "black", "red", "red", "blue", "black", "red", "red", "red", "blue", "red", "red", "grey"))) +
  labs(title = "Total number of monoclonals recognizing an antigen",
       x = "Swine Virus",
       y = "Number of Mabs")

ggsave(filename = here::here("results", "figures", "supp_univariate_antigen.png"), plot = univariate_antigen)

```
The P1 monoclonal antibodies recognized 13 out of 14 viruses, and the CA/09 antibodies recognized 13 out of 14 as well. There was one virus specific to P1 recognition (A/Swine/North Carolina/2009), and one virus specific to just CA/09 recognition (A/Swine/Utah/2009). The A/Swine/North Carolina/5043-1/2009 virus keeps falling within the same groups as the classical lineage viruses to the point where it is worrisome. I double checked the sequence in the lab used to make the virus-like particle and the classficiation of the virus with the literature and it confirmed that it is supposed to be in the human seasonal-like lineage. 
