---
title: "processing_script"
author: "Amanda Skarlupka"
date: "10/26/2019"
output:
  word_document: default
  pdf_document: default
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


load the packages needed

```{r}
library(readxl)
library(knitr)
library(tidyverse)
library(readxl)
library(gtools)
library(gdata)
```


load the data. The data is contained within one excel file, but is separated on different sheets
```{r}

p1_mab_full <- read_excel("~/Documents/Data Science/AmandaSkarlupka-Project/data/raw_data/data.xlsx", 
                      sheet = "p1_mab", na = "NA")
ca09_mab_full <- read_excel("~/Documents/Data Science/AmandaSkarlupka-Project/data/raw_data/data.xlsx", 
                      sheet = "ca09_mab", na = "NA")
p1_sera_full <- read_excel("~/Documents/Data Science/AmandaSkarlupka-Project/data/raw_data/data.xlsx", 
                        sheet = "p1_sera", na = "NA")
ca09_sera_full <- read_excel("~/Documents/Data Science/AmandaSkarlupka-Project/data/raw_data/data.xlsx", 
                        sheet = "ca09_sera", na = "NA")
antigen_key <- read_excel("~/Documents/Data Science/AmandaSkarlupka-Project/data/raw_data/data.xlsx", 
                   sheet = "antigen_key", na = "NA")
antibody_key <- read_excel("~/Documents/Data Science/AmandaSkarlupka-Project/data/raw_data/data.xlsx",
                           sheet = "antibody_key", na = "NA")


```



#Take a look at the data and make sure that it loaded correctly
```{r}

summary(p1_mab_full)
str(p1_mab_full)
str(ca09_mab_full)
str(p1_sera_full)
str(ca09_sera_full)

```


Move the monoclonal antibody concentrations and the sera titers into the same column

```{r}

p1_mab <- p1_mab_full %>%
  gather(key = "antigen", value = "concentration", 2:19)

ca09_mab <- ca09_mab_full %>%
  gather(key = "antigen", value = "concentration", 2:19)

p1_sera <- p1_sera_full %>%
  gather(key = "antigen", value = "titer", 2:19)

ca09_sera <- ca09_sera_full %>%
  gather(key = "antigen", value = "titer", 2:19)

```



The sera was not ran against all the viruses due to a limited supply. So the viruses that were used for sera testing are filtered below.

```{r}

ca09_sera <- ca09_sera %>%
  na.omit()
p1_sera <- p1_sera %>%
  na.omit()

```

The antibody concentrations are in two fold dilutions, the titers will be log2(value)

```{r}

p1_mab$log2 <- log2(p1_mab$concentration)
p1_mab$dilution <- log2(1/(p1_mab$concentration/20))*2

p1_sera$log2 <- log2(p1_sera$titer)


ca09_mab$log2 <- log2(ca09_mab$concentration)
ca09_mab$dilution <- log2(1/(ca09_mab$concentration/20))*2

ca09_sera$log2 <- log2(ca09_sera$titer)

```


#The lineages of the viruses need to be added. The antigen_key contains all the information about the viruses. 

```{r}
#add the antigen key to the sera data for switch out the antigen with the antigen key for ease of script writing

ca09_sera <- ca09_sera %>%
  left_join(antigen_key, by = c("antigen" = "short_name"))

p1_sera <- p1_sera %>%
  left_join(antigen_key, by = c("antigen" = "short_name"))

ca09_mab <- ca09_mab %>%
  left_join(antibody_key, by = "clone_name")

p1_mab <- p1_mab %>%
  left_join(antibody_key, by = "clone_name")
```


save data as RDS

```{r}

saveRDS(ca09_mab, file = "~/Documents/Data Science/AmandaSkarlupka-Project/data/processed_data/ca09mab_processed_data.rds")
saveRDS(ca09_sera, file = "~/Documents/Data Science/AmandaSkarlupka-Project/data/processed_data/ca09sera_processed_data.rds")
saveRDS(p1_mab, file = "~/Documents/Data Science/AmandaSkarlupka-Project/data/processed_data/p1mab_processed_data.rds")
saveRDS(p1_sera, file = "~/Documents/Data Science/AmandaSkarlupka-Project/data/processed_data/p1sera_processed_data.rds")
#saveRDS(p1_sera_full, file = "./data/processed_data/p1sera_full_processed_data.rds")
#saveRDS(CA09_sera_full, file = "./data/processed_data/CA09sera_full_processed_data.rds")
#saveRDS(p1_mabs_full, file = "./data/processed_data/p1mabs_full_processed_data.rds")
#saveRDS(CA09_mabs_full, file = "./data/processed_data/CA09mabs_full_processed_data.rds")

```


